%************************************************
\chapter{Results}\label{ch:results}
%************************************************

\todo{
difficulties in picking data sets
performance results on locating algorithm
}

\section{Data sets}

Unfortunately, \term{political events} are a rather vague concept judging by the contents of the Wikipedia category of the same name\footurl{http://en.wikipedia.org/wiki/Category:Political_events}{2012}{02}{01}.
It includes only one page, \term{No Berlusconi day}, but a numerous sub-categories ranging from \term{Political riots} and \term{Protests} to \term{Political party assemblies}.

\todo{
more homogenous are article groups formed by template embedders,
like Infobox civil conflict.
}

\todo{
acknowledging that there is no ``right'' categorry or template
goal is to find a group of articles
using event definition, together with handpicked categories and templateembedders in the English Wikipedia.
categories contain articles that are handpickked by wikipedians
}

\todo{
shortcomings, not random, statistics derived cannot be representative to all articles treating political events   let alone the all of wikipedia
misses arcticles that are not part of category, template embedders
}
% Military conflict 606/10290


\subsection{By category}

\todo{
categories can go deep pretty quickly (over-categorition) but also have to potential to lead astray (Protest Songs in Protest)
sub-categories the same pages in different tree structure, e.g. protests by country and protests by year.
the same article appears in different depths of the resulting sub-tree.
category list will go only one level deep
Protests and its sub-categories are exhaustive and prime candidate.
they yield: (and combine)
Protests by year % good, 45.5 % of 143
Protests by country
21st-century conflicts % bad, only 10% of 833 articles qualified
2000s riots by year
2010s riots by year
footnote Revolutions categories, e.g. Revolutions by country, contain too many articles about individuals and not events
}
\todo{link to appendix with complete list}


\subsection{By template}

\todo{
info box civil conflict as a reference, arab springs as well as the occupy movement favorite. 
numbers by filtering
second set: info box election, big set 4485 articles
}
\todo{link to appendix with complete list}


\subsection{Political article vs. place article}

\todo{
To compare the date with spatial articles preferred in previous studies, a reference group is introduced
template info box settlement is used by towns all over the world
population is big 261,000+ articles, pick a random sample
compare locating performance, sd metrics.
}
\todo{link to appendix with complete list}



\begin{todos}
    \item for each set, show hypothesis results, values
    \item box charts
\end{todos}

% H1 susceptible to outliers
